%==============================================================================
% Enchan Field Notes (Theoretical Framework)
% Version: \version (Pivot: Inception & Anchor)
% Main file - loads sections from ./sections/
%==============================================================================

\documentclass[11pt,a4paper]{article}

%===========================
% Packages
%===========================
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{mathptmx}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{bm}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{graphicx}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue,
    pdftitle={Enchan Field Notes v0.4.0},
    pdfauthor={Mitsuhiro Kobayashi},
    pdfsubject={Inception World: Topological Defects and Emergent Acceleration Scale},
    pdfkeywords={Enchan, dark matter, RAR, BTFR, topological defect, scalar field}
}

%===========================
% Helpers
%===========================
% chapter-like section with page break
\newcommand{\chap}[1]{\clearpage\section{#1}}

% Dimensionless field symbol (explicitly defined)
\newcommand{\EnchanS}{S}

%---------------------------
% hyperref: avoid PDF-string warnings when math appears in headings
%---------------------------
\pdfstringdefDisableCommands{%
  \def\EnchanS{S}%
  \def\bm#1{#1}%
  \def\mathrm#1{#1}%
  \def\mathcal#1{#1}%
  \def\sqrt#1{sqrt(#1)}%
  \def\ln{ln}%
}

%===========================
% Title / Author
%===========================
\title{\textbf{Enchan Field Notes \version}\\[0.5em]
\large The Inception World: Topological Defects and the Origin of an Acceleration Scale}

\author{Mitsuhiro Kobayashi\\[0.25em]
Tokyo, Japan\\
\texttt{enchan.theory@gmail.com}}

\date{\today}

%===========================
% Document
%===========================
%--- Version Definition ---
\newcommand{\version}{v0.4.0} 
%--------------------------

\begin{document}

\maketitle
\thispagestyle{empty}

%===========================
% Abstract
%===========================
\begin{abstract}
\noindent
\textbf{\version update (structural refinement):}
This version consolidates the framework into a minimal, explicitly testable form.
We formulate a \emph{dimensionless} scalar field $\EnchanS$, interpreted as an effective
time-dilation factor (a compact proxy for gravitational time delay), and examine how
defect-like configurations of $\EnchanS$ anchored by baryonic structure can generate
galaxy-scale regularities.

A central result of \version is the explicit separation of regimes.
At the galaxy scale, we introduce an \emph{environmental pinning} mechanism
(Eq.~6.6.2) in which the effective acceleration scale
$a_{0,\mathrm{eff}}$ is suppressed in deep baryonic potentials via a potential-depth proxy.
This term is constrained directly by cross-validated fits to SPARC data.

At smaller, high-acceleration scales (e.g.\ the Solar System), we show that an additional
suppression may be required to avoid unphysical extrapolation.
This effect is isolated as a separate safeguard (Eq.~6.6.3) and is \emph{not} used in
galaxy calibration.

The overarching aim of \version is not to propose a modification of gravity,
but to organize observed phenomenology into a controlled effective description,
with clear regime boundaries and falsifiable components.
\end{abstract}


%===========================
% Status / Scope box
%===========================
\vspace{1em}
\hrule
\section*{Status and Scope (\version)}
This document presents a \textbf{speculative effective framework} intended to sit
beneath (not replace) the Standard Model and General Relativity.

\begin{itemize}
    \item \textbf{Empirical baseline:}
    The accompanying Python scripts reproduce standard galaxy-scale regularities
    (SPARC-based RAR / BTFR / fixed-parameter rotation-curve predictions)
    as externally checkable benchmarks.
    Verification code is available at:\\
    \url{https://github.com/EnchanTheory/Enchan-Field-Notes}

    \item \textbf{Theoretical proposal:}
    The working hypothesis is that these regularities can be interpreted as signatures
    of defect-like configurations in the field $\EnchanS$, with baryonic structure acting
    as an anchor.

    \item \textbf{Regime separation (\version):}
    Environmental suppression in deep baryonic potentials (Eq.~6.6.2) is treated as
    part of the galaxy-scale phenomenology and is calibrated on galaxy data.
    Additional high-acceleration suppression (Eq.~6.6.3) is introduced only as an
    optional safeguard for Solar-System diagnostics and is not included in galaxy fits.

    \item \textbf{Assumptions:}
    The anchor condition linking baryonic structure to the effective defect scale
    remains a phenomenological ansatz and is explicitly identified as such.

    \item \textbf{Conservativeness:}
    The framework does not address strong-field gravity or precision Solar-System tests;
    its domain of applicability is the weak-field, galaxy-scale regime where the
    empirical relations are defined.
\end{itemize}
\hrule
\vspace{1em}


\clearpage
\tableofcontents
\clearpage

%===========================
% Content Injection
%===========================
\input{sections/01_overview}
\input{sections/02_concept}
\input{sections/03_rotation}
\input{sections/04_field_model}
\input{sections/05_asymptotics}
\input{sections/06_derivation}
\input{sections/07_observation}
\input{sections/99_conclusion}

%===========================
% Bibliography
%===========================
\clearpage
\begin{thebibliography}{99}

\bibitem{Kibble1976}
T.~W.~B. Kibble,
\newblock ``Topology of cosmic domains and strings,''
\newblock {\em Journal of Physics A: Mathematical and General}, 9:1387, 1976.

\bibitem{BarriolaVilenkin1989}
M. Barriola and A. Vilenkin,
\newblock ``Gravitational field of a global monopole,''
\newblock {\em Physical Review Letters}, 63:341, 1989.

\bibitem{McGaugh2016}
S.~S. McGaugh, F.~Lelli, and J.~M. Schombert,
\newblock ``Radial Acceleration Relation in Rotationally Supported Galaxies,''
\newblock {\em Physical Review Letters}, 117:201101, 2016.

\bibitem{Lelli2019}
F. Lelli et al.,
\newblock ``The Baryonic Tully--Fisher Relation for Different Galaxy Types,''
\newblock {\em Monthly Notices of the Royal Astronomical Society}, 484:3267, 2019.

\bibitem{Abbott2021}
R. Abbott et al. (LIGO Scientific Collaboration and Virgo Collaboration),
\newblock ``Open Data from the First and Second Observing Runs of Advanced LIGO and Advanced Virgo,''
\newblock {\em SoftwareX}, 13:100658, 2021.

\bibitem{Dvali2000}
G. Dvali, G. Gabadadze, and M. Porrati,
\newblock ``4D Gravity on a Brane in 5D Minkowski Space,''
\newblock {\em Physics Letters B}, 485:208, 2000.

\end{thebibliography}

\end{document}